cardinalR: Generating interesting high-dimensional data structures

### 2025-10-08 paper-cardinalR.Rmd
Try to make a complex dataset

positions <- geozoo::simplex(p=4)$points
positions <- positions * 0.8

## To generate data
five_clusts <- gen_multicluster(n = c(2250, 1500, 750, 1250, 1750), k = 5,
                       loc = positions,
                       scale = c(0.4, 0.35, 0.3, 1, 0.3),
                       shape = c("helicalspiral", "hemisphere", "unifcube", 
                                 "cone", "gaussian"),
                       rotation = NULL,
                       is_bkg = FALSE)
```

@@

>>>>

positions <- geozoo::simplex(p=4)$points
positions <- positions * 0.3 #positions * 0.8

## To generate data
five_clusts <- gen_multicluster(n = c(2250, 1500, 750, 1250, 1750), k = 5,
                       loc = positions,
                       scale = c(0.25, 0.35, 0.3, 1, 0.3),
                       shape = c("helicalspiral", "hemisphere", "unifcube", 
                                 "cone", "gaussian"),
                       rotation = NULL,
                       is_bkg = FALSE)
```

@@

```{r}
library(fpc)

# Extract high-dimensional data
data <- five_clusts[, -5]
true_labels <- as.numeric(gsub("cluster", "", five_clusts$cluster))

# Hierarchical clustering
dist_mat <- dist(data)
hc_res <- cutree(hclust(dist_mat, method = "ward.D2"), k = 5)
```

```{r}
# K-means clustering
kmeans_res <- kmeans(data, centers = 5, nstart = 20)
```

```{r}
library(mclust)

# Model-based clustering
mclust_res <- Mclust(data, G = 5, modelNames = "VVV")$classification
```

```{r}
# Evaluate clustering performance
comp_kmeans <- cluster.stats(dist_mat, true_labels, kmeans_res$cluster)
comp_hclust <- cluster.stats(dist_mat, true_labels, hc_res)
comp_mclust <- cluster.stats(dist_mat, true_labels, mclust_res)

# Summarize key performance metrics
data.frame(
  Method = c("K-means", "Hierarchical", "Model-based"),
  Avg_Silhouette = c(comp_kmeans$avg.silwidth, comp_hclust$avg.silwidth, comp_mclust$avg.silwidth),
  Dunn_Index = c(comp_kmeans$dunn, comp_hclust$dunn, comp_mclust$dunn),
  Entropy = c(comp_kmeans$entropy, comp_hclust$entropy, comp_mclust$entropy),
  Corrected_Rand = c(comp_kmeans$corrected.rand, comp_hclust$corrected.rand, comp_mclust$corrected.rand)
)

### 2025-10-08 paper-cardinalR.tex
Try to make a complex dataset

\begin{verbatim}
positions <- geozoo::simplex(p=4)$points
positions <- positions * 0.8

## To generate data
five_clusts <- gen_multicluster(n = c(2250, 1500, 750, 1250, 1750), k = 5,
                       loc = positions,
                       scale = c(0.4, 0.35, 0.3, 1, 0.3),
                       shape = c("helicalspiral", "hemisphere", "unifcube", 
                                 "cone", "gaussian"),
                       rotation = NULL,
                       is_bkg = FALSE)
\end{verbatim}

@@

@@

The source material for this paper is available at \href{https://github.com/JayaniLakshika/paper-cardinalR}{github.com/JayaniLakshika/paper-cardinalR}.

This article is created using \CRANpkg{knitr} \citep{yihui2015} and \CRANpkg{rmarkdown} \citep{yihui2018} in R with the \texttt{rjtools::rjournal\_article} template. These \texttt{R} packages were used for this work: \texttt{cli} \citep{gabor2025}, \texttt{tibble} \citep{kirill2023}, \texttt{gtools} \citep{gregory2023}, \texttt{dplyr} \citep{hadley2023}, \texttt{stats} \citep{core2025}, \texttt{tidyr} \citep{hadley2024}, \texttt{purrr} \citep{hadley2025}, \texttt{mvtnorm} \citep{alan2009}, \texttt{geozoo} \citep{barret2016}, and \texttt{MASS} \citep{venables2002}.

>>>>

\begin{verbatim}
positions <- geozoo::simplex(p=4)$points
positions <- positions * 0.3 #positions * 0.8

## To generate data
five_clusts <- gen_multicluster(n = c(2250, 1500, 750, 1250, 1750), k = 5,
                       loc = positions,
                       scale = c(0.25, 0.35, 0.3, 1, 0.3),
                       shape = c("helicalspiral", "hemisphere", "unifcube", 
                                 "cone", "gaussian"),
                       rotation = NULL,
                       is_bkg = FALSE)
\end{verbatim}

@@

\begin{verbatim}
#>         Method Avg_Silhouette  Dunn_Index  Entropy Corrected_Rand
#> 1      K-means       0.227417 0.008968602 1.551532      0.4223504
#> 2 Hierarchical       0.227417 0.008968602 1.551532      0.5029130
#> 3  Model-based       0.227417 0.008968602 1.551532      0.7525318
\end{verbatim}

@@

The source material for this paper is available at \href{https://github.com/JayaniLakshika/paper-cardinalR}{github.com/JayaniLakshika/paper-cardinalR}. This article is created using \CRANpkg{knitr} \citep{yihui2015} and \CRANpkg{rmarkdown} \citep{yihui2018} in R with the \texttt{rjtools::rjournal\_article} template. These \texttt{R} packages were used for this work: \texttt{cli} \citep{gabor2025}, \texttt{tibble} \citep{kirill2023}, \texttt{gtools} \citep{gregory2023}, \texttt{dplyr} \citep{hadley2023}, \texttt{stats} \citep{core2025}, \texttt{tidyr} \citep{hadley2024}, \texttt{purrr} \citep{hadley2025}, \texttt{mvtnorm} \citep{alan2009}, \texttt{geozoo} \citep{barret2016}, and \texttt{MASS} \citep{venables2002}.
